---
title: "Admission_prediction"
author: "Sabrina Mobassirin"
date: "2023-09-24"
output: pdf_document
---

```{r}
library(ggplot2)
library(kableExtra)
library(dplyr)
library(caTools)


```


```{r}
#EDA
#Dataset consists of 400 observations and 8 features.

# data load and remove serual no.
graduate <- read.csv("~/Downloads/Fall23/Stat5301_Applied Regression/Project/Admission/Admission_Predict.csv") %>% dplyr::select(-Serial.No.)
# data overview
glimpse(graduate)
## Rows: 400
```

```{r}
#Change the column name

names(graduate)[1] <- "GRE_Score"
names(graduate)[2] <- "TOEFL_Score" 
names(graduate)[3] <- "University_Rating" 
names(graduate)[8] <- "Admission_Chance"
#Checking for null values.

# null value check
graduate[!complete.cases(graduate),]

#to see relationship between variables
plot(graduate, col="lightblue")
```
```{r}
library(ggplot2)
library(GGally)
ggpairs(graduate,lower = list(continuous = wrap('points', colour = "purple")),
  diag = list(continuous = wrap("barDiag", colour = "violet"))
    )

```
The above plot shows high positive co-relation with GRE score, TOEFL score and CGPA the high chance of getting admission.

```{r}
set.seed(2)

sample = sample.split(graduate$Admission_Chance, SplitRatio = 0.70)

train = subset(graduate, sample == TRUE)
test = subset(graduate, sample == FALSE)

print(dim(train))

print(dim(test))
```

Model building
We will make the multiple linear regression modeing using Chance of Admit as the target variable. We will include all the independent variable.
```{r}
model_mlr <- lm(Admission_Chance ~ ., data = graduate)
summary(model_mlr)
```
 Model showing low standard error, high R2
 and low p-value (< 0.05). To check linearity we will see the residual analysis.
 
```{r}
hist(resid(model_mlr), col = 'skyblue', main = "Residual Disribution", xlab = "Residuals")
```


```{r}
plot(model_mlr)
```
Above residual plot show model met the linearity assumption. So we can write the linear equation:
AdmissionChance=−1.25+0.0017∗GREScore+0.003∗TOEFLScore+0.006∗UniversityRating+−0.003∗SOP+0.022∗LOR+0.119∗CGPA+0.024∗Research

#####
Prediction on test data
To predict the admission chance apply this model on test data. We will add new column named Admission Probability. It shows 2 values 0 and 1, of 0 the probability of admission is less if 1 then probability of admission is high.

```{r}
Predict <- predict(model_mlr, test)
test$Predict <- ifelse(Predict < 0.6, "0", "1")
kable(test[1:10,]) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),latex_options="scale_down")

```
